
\chapter{Background: algorithmic tools for molecule discovery}
\label{chapter:background}

\ifpdf
    \graphicspath{{chapter02-background/figures}}
\else
    \graphicspath{}
\fi

This chapter provides a general overview of existing approaches to molecule discovery.
This is an incredibly large field with an enormous body of prior work
from chemistry, cheminformatics, and machine learning---
far too much to cover exhaustively.
Therefore, this chapter primarily aims to introduce the concepts and methods which
are necessary to understand the later chapters.
A brief summary of the chapter is provided at the end (\S\ref{sec:background:summary}).
Readers who are already familiar with the field may wish to skip directly to this summary
and then proceed to Chapter~\ref{chapter:lso}.

\input{chapter02-background/sections/mols}
\input{chapter02-background/sections/problems}
\input{chapter02-background/sections/models}
\input{chapter02-background/sections/molgen}
\input{chapter02-background/sections/synthesis}
\input{chapter02-background/sections/bayesopt}



\section{Chapter summary and further reading}
\label{sec:background:summary}

This chapter first introduced molecules and their representations (\S\ref{sec:background:mols}).
The space of molecules is denoted $\molspace$, while $\amolspace\subseteq\molspace$ denotes
a space of ``accessible'' molecules.
Molecules are essentially represented as graphs (with a small amount of standardization).
These graphs can be post-processed into classical cheminformatics \emph{descriptors} (e.g.\@ molecular weight, number of rings),
converted into \emph{embeddings} with some kind of neural network,
represented as a string of characters (e.g.\@ SMILES),
or represented as a set of fragments called a \emph{fingerprint}.

Molecule discovery is mainly formalized as an optimisation problem (\S\ref{sec:background:molopt}),
defined in equation~\ref{eqn:defn:molopt}.
Because predicting properties of molecules is clearly useful for optimisation,
we also introduce supervised learning as a sub-problem of optimisation (\S\ref{sec:background:supervised learning}).

The next sections introduced key models and algorithms for molecule discovery.
The most important class is supervised learning models (\S\ref{sec:background:models}),
particularly \emph{Gaussian processes} (GPs, \S\ref{sec:background:gps}).
Generative models for molecules (\S\ref{sec:background:molgen}) were briefly introduced,
as they are used chapter~\ref{chapter:lso}.
Methods to synthesize molecules (\S\ref{sec:background:synthesis}) were also introduced
in preparation for chapter~\ref{chapter:rfb},
with the retro* algorithm being particularly important.

Finally, Bayesian optimisation (\S\ref{sec:background:bayesopt}) was introduced as a general procedure
which uses models to select experiments to perform.
Bayesian optimisation serves as a unifying theme for the rest of the thesis,
as all the methods described in the following chapters (\ref{chapter:lso}--\ref{chapter:rfb})
can be used as part of a Bayesian optimisation procedure.

We now proceed to the main content of this thesis,
starting with an end-to-end BO method for molecule optimisation
in the next chapter.
