
\begin{acknowledgements}      

So many people made my PhD journey possible.
First and foremost I would like to thank my PhD supervisor, Miguel.
I think Miguel strikes a very good balance in his supervision style.
He gave me the freedom to explore my own ideas right from the start of my PhD,
and would always be encouraging, while also being upfront about potential flaws in my ideas.
This made me feel responsible for my own research, yet also supported.
I think this was a very good way to learn, and I am very grateful for this.
I'm also very grateful to my advisor, Adrian Weller, for his support and guidance
when I needed it.

I worked with many great collaborators during my PhD.
Erik Daxberger: I appreciated your perpetual calm.
Wenlin Chen: I appreciated your hard work and critical thinking.
Gregor Simm and Marwin Segler: I appreciated your mentorship and impressive chemistry knowledge.
Sergio Bacallado: I appreciated your mentorship, enthusiasm, patience, and ability to explain things clearly.
Miguel Garc{\'i}a-Orteg{\'o}n: I appreciated your insights into the biology behind drug discovery.
Krzysztof Maziarz: I appreciated your mentorship, impressive coding skills,
and legendary attention to detail when reviewing pull requests during my internship at Microsoft Research.
Sarah Lewis: I appreciated your mentorship and independent thinking.

I'm also very grateful to many people in Cambridge who I worked alongside
and enjoyed many interesting discussions with.
In no particular order:
James Allingham,
Javier {Antor\'an},
Vincent Stimper, Ross Clarke, Elre Oldewage,
Andy Lin, Megan Stanley, Runa Eschenhagen, Juyeon Heo, Tor Fjelde,
John Bronskill, Will Tebbutt, Miri Zilka, Matthew Ashman, David Burt,
Talay Cheema,
Vidhi Lalchand,
Bruno Mlodozeniec,
Adrian Goldwaser,
Robert Pinsler, Puria Radmard, Isaac Reid,
Marine Schimel, Jonathan So, Kenza Tazi, Kris Jensen,
Carl Ek, Carl Rasmussen, Richard Turner, and many others.

I was very lucky to have a great group of friends in Cambridge.
David and Stratis: thanks for always being there for me.
You are the best housemates I could have asked for.
Beno\^it, Risa, Ido, Justin, and many others: thanks for the great times.

I would like to thank the Cambridge Trust and the Canadian Centennial Scholarship Fund for funding my PhD:
I could not have come to Cambridge without this support.

Finally, I could not have done this PhD without the support of those closest to me.
Mom, thanks for always being there for me.
Sam, thanks for always providing a grounding perspective.
And Shirley%
\begin{CJK*}{UTF8}{gbsn}
晓琦, 我亲爱的伴侣,
\end{CJK*}
thanks for your love, patience,
and for the boundless joy we share together.

\end{acknowledgements}
